Plant species traits and growth rates: meta-analysis shows correlations change with plant size as predicted

### 2015-05-14 MS.tex
MS (update)

\documentclass[a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
\usepackage{alltt}
\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{fixltx2e}
\usepackage[default,osfigures,scale=0.95]{opensans}
\usepackage{geometry}
\usepackage{longtable}
\usepackage{authblk}
\usepackage[colorinlistoftodos]{todonotes}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2.5cm}
\linespread{2}
\usepackage{hyperref}

%appendix
\usepackage[title,titletoc,toc]{appendix}

% allows for number referencing, as required by updated ecol letters style
\usepackage{etoolbox}
\newbool{MyRefNumbers}
\booltrue{MyRefNumbers} % comment to remove numbers in reference list

\usepackage{natbib}
\bibliographystyle{ecol_let}
\setcitestyle{authoryear,open={(},close={)}}

\newcommand{\bibstar}{* }

@@

\section*{Cover letter (303 on 300 words)}\label{Proposal}

Editor-in-Chief Ecology letters

Dear Dr Holyoak,

May we please submit for Ecology Letters a manuscript titled “Plant species traits and growth rates: meta-analysis shows correlations change with plant size as predicted”? We would be grateful if you would consider it for publication.

The work is meta-analysis, but we suggest it should be considered as a Letter more so than under the heading of Reviews and Syntheses. It reports original conclusions based on new data analysis, and does not attempt broad review. But of course if you prefer to consider it as a review, then we are in your hands.

\textbf{Novelty statement:}
In a nutshell, this paper sorts out the differences across plant species in potential growth rates. We believe it's a very substantial advance. Growth rates are obviously really important for ecology. They have been seen as a core element of plant ecological strategies going back to Grime and Hunt 1975. Over the past 20 years much ecological strategy research has shifted into "trait ecology", using traits such as SLA and wood density and seed mass as indicators of ecological strategy. Yet empirical reports of the correlations between traits and growth rates have been inconsistent. Our manuscript first spells out reasons why traits should influence growth differently depending what size the individual plant has achieved -- in other words, traits should influence growth trajectories rather than simply making species fast-growing versus slow-growing throughout life. Then second, we provide meta-analysis of correlations that have been reported between traits and growth rate. To the extent data are available, these correlations are shown to shift depending on plant size consistently with the reasons explained and the predictions made in the first part of the paper.

This study is the first one of an upcoming series of theoritical and empirical studies performed by the authors on the same thematic: the effect of plant size on plant strategies and functional traits.

We are excited by this work and look forward to your evaluation of the paper.

Sincerely

Anais Gibert on behalf of all authors

@@

shows predictable changes with plant size of the correlations

@@

\textit{Number of words in the abstract} 148

@@

\textit{Number of words in each text box} 750

@@

\section*{Abstract (148 words/150)}\label{abstract}

@@

\item
Traits should influence growth in a size-dependent manner. We outline mechanisms and hypotheses based on new theory, and test these predictions for five traits using a meta-analysis of 108 studies (\textgreater{} 500 correlations).

@@

\item
We show that correlations between traits and GR change in a predictable way as a function of plant size. Our understanding of plant strategies should shift away from attributing slow vs fast growth to species throughout life, in favour of attributing growth trajectories.

@@

The phrase "plant ecological strategy" refers to how species face the challenge they all share:  to acquire sufficient water, nutrients and light for individual growth and to live long enough to replace themselves with offspring. An idea that strategy differences might involve fast vs slow growth goes back at least to Grime and Hunt \citeyearpar{Grime:1975gr}, who measured a wide spread of relative growth rates for seedlings during early exponential growth under favourable conditions. Seedling potential RGR has continued to be used as a major strategy indicator within the Sheffield CSR (Competitor - Stress tolerator - Ruderal) scheme and its "Integrated Screening Program" (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling RGR into its components showed that specific leaf area (SLA = 1/leaf mass per area LMA) was typically the strongest source of variation between species \citep{Poorter:1989tx}. From the 1990s a new style of "trait ecology" arose whereby measurable traits including SLA were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents and latitudes and thousands of species. Traits have been seen as constructive approaches to understanding plant growth strategies \citep{Grime:1977kc,Chapin:1980gz}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Iida:2014ep,Iida:2014hq,Ruger:2012jv,Falster:2011ii}.Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when  potential influences of traits on plant function are considered \citep{Falster:2011ii, falster:2013} (see Text Box 1 \ref{app:growth} in Supporting Informations). At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover.

Here we review published literature reporting on relationships between
plant growth and five prominent traits (specific leaf area SLA, wood or stem tissue density WD, maximum leaf photosynthetic capacity per unit area Aarea,  maximum height Hmax, and seed mass SM), in order to assess the following questions.

\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions (outlined below) from a recently developed  mechanistic
model \citep{Falster:2011ii, falster:2013}(Text Box 1 \ref{app:growth})?
\end{enumerate}

While most studies on traits have not been concerned with effects of plant stage or size, they have reported trait-growth correlations for identifiable size classes. We gathered correlation coefficients r between trait and growth from studies that gave explicit information about the size or stage of the plants. Meta-analysis allowed us to generalize about trait influences on growth.

@@

Recent theoretical work \citep{Falster:2011ii} allows hypotheses to be formulated about relationships between traits and growth. Below we provide a verbal description for each of the hypotheses; for a mathematical description see Text Box 1 \ref{app:growth}.

\textbf{H1: The relationship between specific leaf area (SLA) and growth rate is expected to shift from positive in seedlings to non-significant or negative in adult plants, regardless of the growth measurement used}

Plants with high SLA can deploy more leaf area per unit of mass invested. However, high-SLA species also have shorter leaf lifespan and faster leaf turnover rate \citep{Wright:2004jb}. Whether plant growth increases with SLA depends on the relative magnitude of the leaf construction and leaf turnover effects of SLA. During early life while leaf comprises a large part of the plant, decreasing leaf construction cost has a positive effect on growth rate. As plants increase in size and leaf mass fraction decreases, the benefit of cheaper leaf construction diminishes; thus the cost of replacing leaf turnover becomes more important.

@@

Low WD decreases the cost of building stem tissue, which should result in faster growth. Although cheaper wood carries a risk of increased mortality, this cost is not reflected in growth outcomes. Thus decreasing WD is predicted to increase growth across the entire size range.

\textbf{H3: The relationship between maximum photosynthetic rate (Aarea) and growth rate is expected to be positive across all plant sizes, at least under high-light conditions}

We expect greater photosynthetic capacity to translate into faster growth rates, at least at high light where Aarea is measured. It is uncertain how much this might translate to faster growth in low light.

\textbf{H4: The relationship between asymptotic maximum height (Hmax) and growth rate is expected to be absent in seedlings but to strengthen with increasing plant size.}

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do, Falster:2011ii, Wenk:2014jz}). For seedlings, where all individuals are investing solely in growth, we expect no influence of Hmax on growth. As plants become larger, species with smaller Hmax are expected to be allocating more to reproduction and less to growth, because at any given size they are closer to their Hmax.

\textbf{H5: The relationship between seed mass (SM) and relative growth rate is negative in seedlings, but weakens with increasing size.}

@@

@@

We aimed to identify studies that measured the correlation between a functional trait and a growth measurement, and searched for relevant studies using a three step procedure. First, we identified traits for which a substantial number of studies would be available, using a literature snapshot in Web of Science. While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are restricted to metadata, such as title, keywords and abstracts \citep{Beckmann:2012hn}. This limitation does not apply to Google Scholar, where, theoretically, all text that is electronically available on the Internet will be indexed. Thus, our second step was to systematically read documents returned by Google scholar for a restricted set of traits. Finally, we identified and recorded other possible data sources using our own knowledge as researchers in the field. (This included adding data from an as-yet unpublished study by Gleason et al). The complete list of studies where we inspected the full text is given in Appendix \ref{app:supp_info_files}.

@@

We focused on an initial list of 20 traits (Fig. \ref{Fig1}): the five for which we had prior expectations, plus 15 others which capture different aspects of plant species strategies. Specific leaf area (SLA), leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and assimilation rate of CO$_2$ (Aarea) are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. Seed mass (SM), and maximum plant height (Hmax) capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. Wood density (WD) and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost \citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf \citep{Lambers:1992bj}. We used the literature snapshot to identify traits which i) were most likely to return a substantial number of articles , and ii) shared a large proportion of their literature list with other traits. The Web of Science database (Thomson Reuters) was searched between March and June 2014 using a range of Boolean search terms in English. Search terms were as follows: (``NAME'' AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''), ``NAME'' was one of trait names listed in Table \ref{TableA1}. According to this snapshot, 62\% of the literature for the 20 traits (or 82\% for the 5 traits for which we have clear predictions) is captured by searching only on two traits: SLA and WD.

@@

We tested size/stage effect on traits for which \textgreater{}20 correlation coefficients across at least two stages were recorded. We used a mixed-effect model of meta-analysis, in which it is assumed that studies within a stage share a common mean effect but that there is also random variation among studies in a stage, in addition to sampling variation. This random effect allows us to account for the heterogeneity among studies (e.g.~experimental type, vegetation type, error, etc.). For each trait, we established if stage could explain variance in the coefficient correlation by comparing models with the stage variable to the null model using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the null model. Effects of other variables (vegetation type, experiment type and growth measurement) were also tested individually (results shown in the Fig. \ref{FigA7}). Estimates of the effect size for each stage were considered significant if the 95\% confidence intervals did not overlap with zero. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and re-calculated the effect size. Finally, we checked the dataset for publication bias with funnel plots (\citealt{Koricheva:2013tz}, Fig. \ref{FigA8}). A publication bias toward significant results creates an asymmetric funnel with small studies with non-significant effects missing from the mouth of the funnel on the side opposite to the true effect. In addition, to avoid publication bias due to selective reporting of significant results within published articles (e.g.~particular case of a `failure to publish bias' see \citealt{Jennions:2013ta}), we asked the authors for additional information when they reported a non-significant correlation between growth and trait without reporting an r value. All meta-analyses were conducted with R software \citep{Ralanguageanden:2014wf}.

@@

Given our basic conclusion that reported correlations responded to plant size in a manner consistent with expectations from theory, we assess now the coverage of the available data, to determine if biased sampling could have erroneously generated the above results. The results reported were confined to correlations across species and under non-stressed conditions. A "complete" dataset (n=328 vs n=198) also included correlations measured at within-species level, and under stressed conditions, and spanning different plant stages. This "complete" dataset generated similar effect sizes to those reported above from the "ideal" dataset, and led to similar conclusions (Fig. \ref{Fig2} grey points). Differences concerned often the confidence intervals around particular mean effect sizes in Fig. 2. These differences in CI were not easily attributed to any single cause. For instance, the correlation between growth and WD at juvenile stage shifted from negative to non significant (i.e CI95\% included zero) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct};  these correlations unlike others used wood density values from adult plants together with growth from juveniles. For Aarea, the comparison between complete and ideal data showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 7 correlations measured under low light conditions were added.

Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{FigA6}) .

@@

For WD the meta-analysis showed negative correlation with growth across all plant stages, as previously reported from individual studies \citep{Wright:2010tp,Ruger:2012jv}. It should be noted however that there are few tests of this proposition at seedling stage. In our model (Text Box 1 \ref{app:growth}) cheaper stem construction is responsible for the predicted negative correlation. Lower xylem conductance and associated lower photosynthetic potential have also been suggested \citep[reviewed by][]{Chave:2009iy}.

For the correlation between Aarea and growth, to our knowledge no study has previously assessed this in relation to plant size. In the growth model (Text Box 1 \ref{app:growth}), a consistent correlation irrespective of size arises from a direct effect of Aarea on plant growth rate. In juvenile stage, where most of the data were, we noted no signal of an effect of low light on the positive correlation between Aarea and growth rate.

@@

\section*{Acknowledgements}\label{Acknowledgment}

We thank Andrea Stephens for helpful comments on the meta-analyses, and Sean Gleason et al for pre-publication access to their data. Charles Warren and Jordi Martìnez-Vilalta kindly provided additional information about their published results. This work was funded by the Australian Research Council through a fellowship to Westoby.

@@

\nocite{*}
\bibliography{output/refs}\label{references}

@@

\begin{figure}[htbp]
\centering
\includegraphics{output/Fig2.pdf}
\caption{\textbf{The relationship between growth rates and SLA or seed mass changes with plant stage.} Plots show the effect size (+ 95\%CI) of the correlation between growth rate and five functional traits for plant stage comparisons. \textbf{a)} Specific leaf area (SLA), \textbf{b)} Wood density (WD), \textbf{c)} Asymptotic height (Hmax) \textbf{d)} Seed mass and \textbf{e)} Assimilation rate of CO2 (Aarea). Effect size is a standardized measure of the deviation of correlation coefficient from zero (Fisher's z-transformed). Effects are significant if confidence intervals (95\%CI) for a given stage do not overlap with zero. Black point: ideal data (at interspecific level \& under unstressed conditions), grey points: all the data. 'n' indicates number of correlations contributing to each effect. Likelihood Ratio Tests (LRT, for ideal data only) compared maximum-likelihood fits between a pooled model and one separated by stage. A significant p-value means the stage model fitted the data better than the null model.
}
\label{Fig2}

% Is it Ecol Lett style to have a declarative first sentence in Fig legends stating the main result? -- or should the explanation be more deadpan, relying on the main text to make the point?

\end{figure}

@@

\begin{table}[ht]
\centering
\caption{\textbf{ Hypothesised effect of traits on components of growth.} Arrows indicate the effect an increase in trait value would have on the elements of eqs. \ref{eq:eq0}-\ref{eq:eq4}.}
  \begin{tabular}[c]{lcccc}
  & $\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}$
  & $\frac{\textrm{d}m_\textrm{t}}{\textrm{d}b}$
  & $\frac{\textrm{d}b}{\textrm{d}t}$
  & $k_s$ \\\hline
  Seed size & & & & \\ 
  Aarea & & & $\uparrow$ & \\
  SLA &$\uparrow$ & & $\downarrow$& \\
  Wood density & $\downarrow$ &  & & \\
  Maximum height & &$\uparrow$ & & \\
\hline
  \end{tabular}
\label{tab:trade-offs}
\end{table}

\newpage
\section*{Text Box} \label{app:growth}

@@

We adapted growth equations from \cite{Falster:2011ii} and \cite{falster:2013} to generate hypotheses on both absolute and relative growth rates in biomass ($m_t$, eqn \ref{eq:eq1}), height (H, eqn \ref{eq:eq2}) and diameter (D, eqn \ref{eq:eq4}). These size dimensions are routinely measured on seedling and adult plants.

Let $m_i$ denote the mass, $r_i$ the mass-specific respiration rate, and $k_i$ the turnover rate of different plant tissues, with the subscripts $i=l,s,r,t$ referring to live biomass in leaves, sapwood and bark, roots, and in the plant. Similarly let $a_i$ denote areas, of leaves ($i=l$) and of cross-sections of total stem and sapwood ($i= st,ss$) respectively.

The rate at which live biomass originating from the plant $b$ changes is given by
\begin{equation}\label{eq:eq0}
\underbrace{\frac{db}{dt}}_{\text{net live biomass production}} = \underbrace{y}_{\text{yield}}  \big(\underbrace{a_{l} A_{\textrm{area}}}_{\text{photosynthesis}} - \underbrace{\sum_{i=l,s,r} m_{i} r_{i}}_{\text{respiration}}\big) - \underbrace{ \sum_{i=l,s,r} m_{i} k_{i}}_{\text{turnover}},
\end{equation}
where $y$ is the fraction of assimilated carbon fixed into biomass,  and  $A_{\textrm{area}}$ is the assimilation rate of CO$_{2}$ per leaf area. The rate at which live biomass is accumulated on the plant is then

>>>>

\documentclass[a4paper]{article}\usepackage[]{graphicx}\usepackage[]{color}
\usepackage{alltt}
\usepackage{amssymb,amsmath}
\usepackage{palatino}
\usepackage{framed}
\usepackage{parskip}
\usepackage{graphicx}
\usepackage{fixltx2e}
\usepackage[default,osfigures,scale=0.95]{opensans}
\usepackage{geometry}
\usepackage{longtable}
\usepackage{authblk}
\usepackage[colorinlistoftodos]{todonotes}
\geometry{verbose,tmargin=2cm,bmargin=2cm,lmargin=2cm,rmargin=2.5cm}
\linespread{2}
\usepackage{color}

\usepackage[title,titletoc,toc]{appendix}

\usepackage{etoolbox}
\newbool{MyRefNumbers}
\booltrue{MyRefNumbers}

\usepackage{natbib}

\bibliographystyle{ecol_let}
\setcitestyle{authoryear,open={(},close={)}}
\newcommand{\bibstar}{* }

@@

@@

@@

\textit{Number of words in the abstract} 150

@@

\textit{Number of words in each text box} 770

@@

\section*{Abstract}\label{abstract}

@@

\item 
Traits should influence growth in a way that depend on individual plants size. We outline mechanisms and hypotheses based on new theory, and test these predictions for five traits using a meta-analysis of 108 studies (\textgreater{} 500 correlations).

@@

\item 
We show that traits-GR correlations change in a predictable way as a function of plant size. Our understanding of plant strategies should shift away from attributing slow vs fast growth to species throughout life, in favour of attributing growth trajectories.

@@

The phrase "plant ecological strategy" refers to how species face the challenge they all share:  to acquire sufficient water, nutrients and light for individual growth and to live long enough to replace themselves with offspring. An idea that species strategy differences might involve fast vs slow growth goes back at least to Grime and Hunt \citeyearpar{Grime:1975gr}, who measured a wide spread of relative growth rates for seedlings during early exponential growth under favourable conditions. Seedling potential RGR has continued to be used as a major strategy indicator within the Sheffield CSR (Competitor - Stress tolerator - Ruderal) scheme and its "Integrated Screening Program" (e.g. \citealt{grime1979plant, Grime:1997wm}). Meanwhile partitioning of seedling RGR into its components showed that specific leaf area (SLA = 1/leaf mass per area LMA) was typically the strongest source of variation between species \citep{Poorter:1989tx}. From the 1990s a new style of "trait ecology" arose whereby measurable traits including SLA were used directly as strategy axes \citep{Westoby:2002ft}. An approach via measurable traits made broad comparisons possible across continents and latitudes and thousands of species. Traits have been seen as constructive approaches to understanding plant growth strategies \citep{Grime:1977kc,Chapin:1980gz}, community composition \citep{Lavorel:2002ff,Shipley:2006ie}, global vegetation dynamics \citep{Scheiter:2013ed} and ecosystem processes \citep{Lavorel:2002ff}.

@@

Meanwhile the possibility has been raised that trait-growth correlations may change systematically with size \citep{Falster:2011ii, Ruger:2012jv, Iida:2014ep, Iida:2014hq}. Recent theory suggests that the absence of a correlation between SLA and growth at larger sizes may be expected, when  potential influences of traits on plant function are considered \citep{Falster:2011ii, falster:2013}. At larger sizes the low-cost deployment benefits of high SLA may be offset or even outweighed by the high costs of rapid leaf turnover (see Text Box 1\ref{sec:growth}).

Here, we review published literature reporting on relationships between
plant growth and five prominent traits (specific leaf area = SLA, wood or stem tissue density = WD, maximum leaf photosynthetic capacity per unit area = Aarea,  maximum height= Hmax, and seed mass= SM), in order to assess the following questions.

\begin{enumerate}
  \item How general is the tendency for correlations between traits and growth rate to vary with plant size?
  \item Do the observed relationships conform to predictions (outlined below) from a recently developed  mechanistic
model \citep{Falster:2011ii, falster:2013}?
\end{enumerate}

While most studies on traits have not been concerned with effects of plant stage or size, they have reported trait-growth correlations for identifiable size classes. We gathered correlation coefficients between trait and growth from studies that gave explicit information about the size or stage of the plants. Meta-analysis allowed us to generalize about trait influences on growth.

@@

Recent theoretical work \citep{Falster:2011ii} allows hypotheses to be formulated about relationships between traits and growth, based on a mathematical decomposition of plant growth rate. Specific predictions arise from considering the influence of traits on elements of these equations and how these effects may vary with size (see Text Box 1\ref{sec:growth}).

\textbf{H1: The relationship between specific leaf area (SLA) and growth rate is expected to shift from positive in seedlings to non-significant or negative in adult plants}

Plants with high SLA can deploy more leaf area per unit of mass invested. However, high-SLA species also have shorter leaf lifespan and faster leaf turnover rate \citep{Wright:2004jb}. Whether plant growth increases with SLA depends on the relative magnitude of the leaf construction and leaf turnover effects of SLA (Table 1). During early life, while leaf comprises a large part of the plant, decreasing leaf construction cost has a positive effect on growth rate. As plants increase in size and leaf mass fraction decreases, the benefit of cheaper leaf construction diminishes; thus the cost of replacing leaf turnover becomes more important.

@@

Low WD decreases the cost of building stem tissue and thus the cost of deploying additional leaf area (Table 1), which should result in faster growth. Although cheaper wood carries a risk of increased mortality, this cost is not reflected in growth outcomes. Thus decreasing WD is predicted to increase growth across the entire size range.

\textbf{H3: The relationship between asymptotic maximum height (Hmax) and growth rate is expected to be absent in seedlings but to strengthen with increasing plant size.}

As plants approach maximum height they increasingly divert energy away from growth and into reproduction (see \citealt{Thomas:1996do, Falster:2011ii, Wenk:2014jz}). For seedlings, where all individuals are investing solely in growth, we expect no influence of Hmax on growth. As plants become larger, species with smaller Hmax are expected to be allocating more to reproduction and less to growth, because at any given size they are closer to their Hmax (Table 1).

\textbf{H4: The relationship between seed mass (SM) and relative growth rate is negative in seedlings, but weakens with increasing size.}

@@

\textbf{H5: The relationship between maximum photosynthetic rate (Aarea) and growth rate is expected to be positive across all plant sizes, at least under high-light conditions}

We expect greater photosynthetic capacity to translate into faster growth rates by increasing the rate of biomass production (Table 1), at least at high light where Aarea is measured. It is uncertain how much this might translate to faster growth in low light.

@@

We aimed to identify studies that measured the correlation between a functional trait and a growth measurement, and searched for relevant studies using a three step procedure. First, we identified traits for which a substantial number of studies would be available, using a literature snapshot in Web of Science. While Web of Science guarantees a relatively stable and reproducible search environment, with clearly defined lists of indexed journals, its searches are restricted to metadata, such as title, keywords and abstracts \citep{Beckmann:2012hn}. This limitation does not apply to Google Scholar, where, theoretically, all text that is electronically available on the Internet will be indexed. Thus, our second step was to systematically read documents returned by Google scholar for a restricted set of traits. Finally, we identified and recorded other possible data sources using our own knowledge as researchers in the field. (This included adding data from an as-yet unpublished study by Gleason et al). The complete list of studies where we inspected the full text is given in Appendix \ref{app:supp_info_files} in Supplementary material.

@@

We focused on an initial list of 20 traits (Fig. \ref{Fig1}): the five for which we had prior expectations, plus 15 others which capture different aspects of plant species strategies. Specific leaf area (SLA), leaf area, leaf nitrogen and phosphorus concentration, net (mass) assimilation rates (NAR), and assimilation rate of CO$_2$ (Aarea) are traits involved in species leaf economic strategy \citep{Wright:2004jb,Wright:2010tp}. Seed mass (SM), and maximum plant height (Hmax) capture information about reproductive strategy \citep{Falster:2005bw,Moles:2006ft}. Wood density (WD) and sapwood area per leaf area (SA/LA), vessel size and density, and stem specific conductance are all involved in species wood economic strategy \citep{Chave:2009iy}. Here WD and SA/LA are also considered as indicators of stem construction cost \citep{Falster:2011ii}, whereas SLA and leaf thickness capture crucial information about leaf construction cost \citep{Wright:2004jb}. Finally, the leaf area ratio and leaf mass ratio are indicators of resource allocation to leaf \citep{Lambers:1992bj}. We used the literature snapshot to identify traits which i) were most likely to return a substantial number of articles , and ii) shared a large proportion of their literature list with other traits. The Web of Science database (Thomson Reuters) was searched between March and June 2014 using a range of Boolean search terms in English. Search terms were as follows: (``NAME'' AND (``growth rate'' OR ``relative growth rate'')) AND (``tree*'' OR ``woody*''), ``NAME'' was one of trait names listed in

\ref{TableA1}. According to this snapshot, 62\% of the literature for the 20 traits (or 82\% for the 5 traits for which we have clear predictions) is captured by searching only on two traits: SLA and WD.

@@

We tested size/stage effect on traits for which \textgreater{}20 correlation coefficients across at least two stages were recorded. We used a mixed-effect model of meta-analysis, in which it is assumed that studies within a stage share a common mean effect but that there is also random variation among studies in a stage, in addition to sampling variation. This random effect allows us to account for the heterogeneity among studies (e.g.~experimental type, vegetation type, error, etc.). For each trait, we established if stage could explain variance in the coefficient correlation by comparing models with the stage variable to the null model using log-likelihood ratios \citep{Zuur:2009cfa}. If the log-likelihood test returned a significant p-value, the stage model fitted the data better than the null model. Effects of other variables (vegetation type, experiment type and growth measurement) were also tested individually (results shown in the Fig. \ref{FigA7}). Estimates of the effect size for each stage were considered significant if the 95\% confidence intervals did not overlap with zero. In the case of multiple comparisons within a study, we used the method proposed by \citet{Borenstein:2009um}: we calculated the mean correlation coefficient by trait and by study, and re-calculated the effect size. Finally, we checked the dataset for publication bias with funnel plots (\citealt{Koricheva:2013tz}, Fig. \ref{FigA8}). A publication bias toward significant results creates an asymmetric funnel with small studies with non-significant effects missing from the mouth of the funnel on the side opposite to the true effect. In addition, to avoid publication bias due to selective reporting of significant results within published articles (e.g.~particular case of a `failure to publish bias' see \citealt{Jennions:2013ta}), we asked the authors for additional information when they reported a non-significant correlation between growth and trait without reporting an r value.

\subsection*{Availability of code and data}\label{code}

All analyses were conducted with R software \citep{Ralanguageanden:2014wf}. The code and data underpinning this paper is available at github.com/AnaisGibert/Growth\_trait\_metaanalysis. Archival copies are also included in Appendix \ref{app:supp_info_files} in Supplementary material.

@@

Given our basic conclusion that reported correlations responded to plant size in a manner consistent with expectations from theory, we assess now the coverage of the available data, to determine if biased sampling could have erroneously generated the above results. The results reported were confined to correlations across species and under non-stressed conditions. A "complete" dataset (n=328 vs n=198) also included correlations measured at within-species level, and under stressed conditions, and spanning different plant stages. This "complete" dataset generated similar effect sizes to those reported above from the "ideal" dataset, and led to similar conclusions (Fig. \ref{Fig2} grey points). Differences concerned often the confidence intervals around particular mean effect sizes in Fig. 2. These differences in CI were not easily attributed to any single cause. For instance, the correlation between growth and WD at juvenile stage shifted from negative to non significant (i.e CI95\% included zero) due to 2 positive correlation coefficients from \citet{Augspurger:1984ct};  these correlations unlike others used wood density values from adult plants together with growth from juveniles. For Aarea, the comparison between complete and ideal data showed no signal of an effect of stress (here mainly low light) on the correlation with growth at seedling stage; the correlation was still positive between Aarea and growth even when 7 correlations measured under low light conditions were added. Finally, our results were unlikely to be biased by dominance of one or a few studies that measured a large number of responses, since when multiple responses were condensed into a single average within a trait and within a study, outcomes were very similar (Fig. \ref{FigA6}) .

@@

For WD the meta-analysis showed negative correlation with growth across all plant stages, as previously reported from individual studies \citep{Wright:2010tp,Ruger:2012jv}. It should be noted however that there are few tests of this proposition at seedling stage. In our model (Text Box 1 \ref{sec:growth}) cheaper stem construction is responsible for the predicted negative correlation. Lower xylem conductance and associated lower photosynthetic potential have also been suggested \citep[reviewed by][]{Chave:2009iy}.

For the correlation between Aarea and growth, to our knowledge no study has previously assessed this in relation to plant size. In the growth model (Text Box 1 \ref{sec:growth}), a consistent correlation irrespective of size arises from a direct effect of Aarea on plant growth rate. In juvenile stage, where most of the data were, we noted no signal of an effect of low light on the positive correlation between Aarea and growth rate.

@@

\section*{Acknowledgments}\label{Acknowledgment}

We thank Andrea Stephens for helpful comments on the meta-analyses, and Sean Gleason et al for pre-publication access to their data. Charles Warren and Jordi Martìnez-Vilalta kindly provided additional information about their published results. This work was funded by the Australian Research Council through a fellowship to Westoby and a discovery grant to Falster.

@@

\nocite{*}

\bibliography{output/refs}\label{references}

@@

\begin{figure}[htbp]
\centering
\includegraphics{output/Fig2.pdf}
\caption{\textbf{The relationship between growth rates and SLA or seed mass changes with plant stage.} Plots show the effect size (+ 95\%CI) of the correlation between growth rate and five functional traits for plant stage comparisons. \textbf{a)} Specific leaf area (SLA), \textbf{b)} Wood density (WD), \textbf{c)} Asymptotic height (Hmax) \textbf{d)} Seed mass and \textbf{e)} Assimilation rate of CO2 (Aarea). Effect size is a standardized measure of the deviation of correlation coefficient from zero (Fisher's z-transformed). Effects are significant if confidence intervals (95\%CI) for a given stage do not overlap with zero. Black point: ideal data (at interspecific level \& under unstressed conditions), grey points: all the data. 'n' indicates number of correlations contributing to each effect. Likelihood Ratio Tests (LRT, for ideal data only) compared maximum-likelihood fits between a pooled model and one separated by stage. A significant p-value means the stage model fitted the data better than the null model.
}
\label{Fig2}
\end{figure}

@@

\begin{table}[ht]
\centering
\caption{\textbf{ Hypothesised effects of traits on key elements of plant function determining growth rate.} Arrows indicate the effect an increase in trait value would have on the main elements of growth eqs. \ref{eq:eq0}-\ref{eq:eq4} in Text box 1:  Leaf deployment per mass ($\frac{\textrm{d}a_\textrm{l}}{\textrm{d}m_\textrm{t}}$), Allocation to vegetative growth ($\frac{\textrm{d}m_\textrm{t}}{\textrm{d}b}$), Net live biomass production ($\frac{\textrm{d}b}{\textrm{d}t}$), Sapwood turnover ($k_s a_s$). SLA: Specific leaf area, WD: Wood density, Hmax: Asymptotic maximum height, SM: seed mass, Aarea: Maximum photosynthetic rate. For further details, see Text Box 1.}
\vspace{1cm}
  \begin{tabular}{l p{3cm} p{2cm}p{4cm} p{2cm}}
  & Leaf deployment \newline per mass
  & Allocation to vegetative growth
  & Net live biomass \newline production
  & Sapwood turnover\\
  & & & &\\\hline
  & & & & \\
Increase in : & & & & \\ 
  SLA &$\nearrow$ &$\_$ & $\searrow$ \newline (via $\nearrow$ of tissue turnover) &$\_$ \\
  WD & $\searrow$ &$\_$  &$\_$ &$\_$ \\
  Hmax &$\_$ &$\nearrow$ &$\_$ &$\_$ \\
  SM &$\_$&$\_$&$\_$& $\_$\\ 
  Aarea &$\_$ &$\_$ & $\nearrow$ &$\_$ \\

\hline
  \end{tabular}
\label{tab:trade-offs}
\end{table}

\newpage
\section*{Text Box 1} \label{sec:growth}

@@

We adapted growth equations from \cite{Falster:2011ii} and \cite{falster:2013} to generate hypotheses on both absolute and relative growth rates in mass (eqn \ref{eq:eq1}), height (eqn \ref{eq:eq2}) and diameter (eqn \ref{eq:eq4}).

The rate at which \textbf{live biomass} $b$ is produced by a plant is given by:

\begin{equation}\label{eq:eq0}
\underbrace{\color{green}\frac{db}{dt}}_{\text{net live biomass production}} = \underbrace{y}_{\text{yield}}  \big(\underbrace{a_{l} A_{\textrm{area}}}_{\text{photosynthesis}} - \underbrace{\sum_{i=l,s,r,f} m_{i} r_{i}}_{\text{respiration}}\big) - \underbrace{ \sum_{i=l,s,r,f} m_{i} k_{i}}_{\text{turnover}},
\end{equation}

where $y$ is the ratio of carbon fixed in mass per carbon assimilated,  and  $A_{\textrm{area}}$ is the assimilation rate of CO$_{2}$ per leaf area.

Here, and in the following equations, $m_i$ denotes the \textbf{live mass on the plant}, $r_i$ the mass-specific \textbf{respiration rate}, and $k_i$ the \textbf{turnover rate} of different plant tissues; the subscripts $i=l,s,r,t,f$ referring to leaves, sapwood and bark, roots, total of these vegetative tissues, and to the reproductive tissues (fruits, flowers etc.) (note that the live biomass $b= m_t +m_f$). Similarly $a_i$ denotes \textbf{areas}, of leaves ($i=l$) and of cross-sections of total stem and sapwood ($i= st,ss$) respectively.

The \textbf{growth rate in mass}, being the increment in vegetative live mass ($m_t$), is then:

\begin{equation}\label{eq:eq1}
\underbrace{{\color{magenta}\frac{dm_t}{dt}}}_{\text{growth rate in mass}} = \underbrace{\color{blue}\frac{dm_t}{db}}_{\text{allocation to vegetative growth}}  \times \underbrace{\color{green}\frac{db}{dt}}_{\text{net live biomass production}},
\end{equation}

where allocation to growth is the fraction of net live biomass increase that goes to vegetative growth rather than to reproductive tissue production.

The \textbf{growth rate in height} can then be expressed as:
